### 2018-01-29 submission/manuscript.Rmd
nick discussion additions

Our observations build upon previous studies [@taq_Gohl_2016] by showing that even different HiFi DNA polymerase have significant differences in the number of OTUs and that the changes to total OTUs correlate with chimeras not removed after sequence processing [Figure 1-2 & 5]. This is important since many diversity metrics rely on the total number of OTUs as part of their calculations and changes to the total number of OTUs could drastically change the results, as well as the findings. Although the attention has mostly been focused on standardizing and improving collection and extraction methods [@contamination_Salter2014] our observations show that independent of this consideration HiFi DNA polymerase can have a noticeable affect on the OTUs generated that can be found across sub-sampling depth and PCR cycle number [Figure 2-4]. These differences were observed in high biomass samples, where biases introduced by such components like kit contamination may have less of an effect, suggesting that these differences may be exacerbated in low biomass samples.

Although we did not observe strong differences, based on cycle number, using the Bray-Curtis index the data suggests that there may be differences between 15 cycles and higher cycle numbers, such as 30x, that are commonly used. Additionally, there was no difference within individuals between corresponding 5 cycle increments (e.g. 15 to 20, 20 to 25, etc.). Conversely, this may be due to low power and on observation there does seem to be a trend that 20x and 25x communities are very different from each other [Figure 3]. This finding, in conjunction with the PERMANOVA results, suggest that cycle number can change bacterial community calculations but that these differences are minimal once 25 cylces are reached. Increasing the sub-sampling depth, for some DNA polymerase, may reduce some of these observed community differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence differences between the different HiFi DNA polymerases [Figure 5]. The chimera prevalence was strongly correlated with the number of OTUs and this value is relied upon heavily for many different downstream community metric calculations. However, Bray-Curtis analysis with PERMANOVA showed few differences based on DNA polymerase. Since it is possible that many of the increased number of OTUs, generated as cycle number increases, are not highly abundant allowing the Bray-Curtis index to be able to successfully downweight these respective OTUs [@bc_index_Minchin1987]. So, choice of downstream diversity metric could be an important consideration in helping to mitigate these observed changes due to high chimera prevalence in HiFi DNA polymerase.

Our observations suggest that there are clear HiFi dependent differences in both per base error rate and chimeras that cannot be removed using bioinformatic approaches [Figure 4 & 5]. Although it may be a natural assumption that the variation may be due to the DNA polymerase family, the highest chimera rate, from Platinum, was a family A polymerase while the lowest, from Accuprime, was also an A polymerase [@polymerase_Ishino_2014]. In fact, from the material safety data sheet (MSDS), it is not clear what the difference between the two different mixes really is. Both Accuprime and Platinum contain a recombinant *Taq* DNA polymerase, a *Pycrococcus* spp GB-D polymerase and a platinum *Taq* antibody. It is possible that differences in how the recombinant *Taq* was generated could be the main reason for the differences in chimera rate since all samples were also sequenced at the same time as well as amplified using the same machine.

>>>>

Our observations build upon previous studies [@taq_Gohl_2016] by showing that even different HiFi DNA polymerases have significant differences in the number of OTUs and that changes in total OTUs correlate with chimeras not removed after sequence processing [Figure 1-2 & 5]. This is important since many diversity metrics rely on the total number of OTUs as part of their calculations and changes to the total number of OTUs could drastically change the results, as well as the findings. Our observations show that HiFi DNA polymerase can have a noticeable affect on the OTUs generated and these differences are consistent across sub-sampling depth and PCR cycle number [Figure 2-4]. These differences were observed in high biomass samples, where biases introduced by such components like kit contamination may have less of an effect, suggesting that these differences may be exacerbated in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index, based on cycle number, the data suggests that there may be differences between 15 and 20 cycles versus higher cycle numbers, such as 30x, that are commonly used. Additionally, there was no difference within individuals between corresponding 5 cycle increments (e.g. 15 to 20, 20 to 25, etc.). Conversely, we may not have enough samples to adequately judge the true magnitude difference between 20x and 25x communities and higher cycle numbers analyzed, even though there is a clear trend to suggest that 20x is very different then 25x [Figure 3]. This finding, in conjunction with the PERMANOVA results, suggest that cycle number can change bacterial community calculations but that these differences are minimal once 25 cylces are reached. Increasing the sub-sampling depth, for some DNA polymerase, may reduce some of these observed community differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence differences between the different HiFi DNA polymerases [Figure 5]. The chimera prevalence was strongly correlated with the number of OTUs which is relied upon heavily for many community metric calculations. However, Bray-Curtis analysis with PERMANOVA showed few differences based on DNA polymerase. It is possible that many of the increased number of OTUs are not highly abundant allowing the Bray-Curtis index to be able to successfully downweight these respective OTUs [@bc_index_Minchin1987]. The choice of downstream diversity metric could be an important consideration in helping to mitigate these observed changes due to high chimera prevalence in specific HiFi DNA polymerases (e.g. Platinum).

Our observations suggest that there are clear HiFi DNA polymerase dependent differences in both per base error rate and chimeras that cannot be removed using bioinformatic approaches [Figure 4 & 5]. Although it may be a natural assumption that the variation may be due to the DNA polymerase family, the highest chimera rate, from Platinum, was a family A polymerase while the lowest, from Accuprime, was also an A polymerase [@polymerase_Ishino_2014]. In fact, from the material safety data sheet (MSDS), the differences between the two mixes is not immediately apparent. Both Accuprime and Platinum contain a recombinant *Taq* DNA polymerase, a *Pycrococcus* spp GB-D polymerase and a platinum *Taq* antibody. It is possible that differences in how the recombinant *Taq* was generated could be the main reason for the differences in chimera rate.

### 2018-01-29 submission/manuscript.tex
nick discussion additions

Our observations build upon previous studies (Gohl et al., 2016) by
showing that even different HiFi DNA polymerase have significant
differences in the number of OTUs and that the changes to total OTUs
correlate with chimeras not removed after sequence processing {[}Figure
1-2 \& 5{]}. This is important since many diversity metrics rely on the
total number of OTUs as part of their calculations and changes to the
total number of OTUs could drastically change the results, as well as
the findings. Although the attention has mostly been focused on
standardizing and improving collection and extraction methods (Salter et
al., 2014) our observations show that independent of this consideration
HiFi DNA polymerase can have a noticeable affect on the OTUs generated
that can be found across sub-sampling depth and PCR cycle number
{[}Figure 2-4{]}. These differences were observed in high biomass
samples, where biases introduced by such components like kit
contamination may have less of an effect, suggesting that these
differences may be exacerbated in low biomass samples.

Although we did not observe strong differences, based on cycle number,
using the Bray-Curtis index the data suggests that there may be
differences between 15 cycles and higher cycle numbers, such as 30x,
that are commonly used. Additionally, there was no difference within
individuals between corresponding 5 cycle increments (e.g.~15 to 20, 20
to 25, etc.). Conversely, this may be due to low power and on
observation there does seem to be a trend that 20x and 25x communities
are very different from each other {[}Figure 3{]}. This finding, in
conjunction with the PERMANOVA results, suggest that cycle number can
change bacterial community calculations but that these differences are
minimal once 25 cylces are reached. Increasing the sub-sampling depth,
for some DNA polymerase, may reduce some of these observed community
differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence
differences between the different HiFi DNA polymerases {[}Figure 5{]}.
The chimera prevalence was strongly correlated with the number of OTUs
and this value is relied upon heavily for many different downstream
community metric calculations. However, Bray-Curtis analysis with
PERMANOVA showed few differences based on DNA polymerase. Since it is
possible that many of the increased number of OTUs, generated as cycle
number increases, are not highly abundant allowing the Bray-Curtis index
to be able to successfully downweight these respective OTUs (Minchin,
1987). So, choice of downstream diversity metric could be an important
consideration in helping to mitigate these observed changes due to high
chimera prevalence in HiFi DNA polymerase.

Our observations suggest that there are clear HiFi dependent differences
in both per base error rate and chimeras that cannot be removed using
bioinformatic approaches {[}Figure 4 \& 5{]}. Although it may be a
natural assumption that the variation may be due to the DNA polymerase
family, the highest chimera rate, from Platinum, was a family A
polymerase while the lowest, from Accuprime, was also an A polymerase
(Ishino \& Ishino, 2014). In fact, from the material safety data sheet
(MSDS), it is not clear what the difference between the two different
mixes really is. Both Accuprime and Platinum contain a recombinant
\emph{Taq} DNA polymerase, a \emph{Pycrococcus} spp GB-D polymerase and
a platinum \emph{Taq} antibody. It is possible that differences in how
the recombinant \emph{Taq} was generated could be the main reason for
the differences in chimera rate since all samples were also sequenced at
the same time as well as amplified using the same machine.

>>>>

Our observations build upon previous studies (Gohl et al., 2016) by
showing that even different HiFi DNA polymerases have significant
differences in the number of OTUs and that changes in total OTUs
correlate with chimeras not removed after sequence processing {[}Figure
1-2 \& 5{]}. This is important since many diversity metrics rely on the
total number of OTUs as part of their calculations and changes to the
total number of OTUs could drastically change the results, as well as
the findings. Our observations show that HiFi DNA polymerase can have a
noticeable affect on the OTUs generated and these differences are
consistent across sub-sampling depth and PCR cycle number {[}Figure
2-4{]}. These differences were observed in high biomass samples, where
biases introduced by such components like kit contamination may have
less of an effect, suggesting that these differences may be exacerbated
in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index,
based on cycle number, the data suggests that there may be differences
between 15 and 20 cycles versus higher cycle numbers, such as 30x, that
are commonly used. Additionally, there was no difference within
individuals between corresponding 5 cycle increments (e.g.~15 to 20, 20
to 25, etc.). Conversely, we may not have enough samples to adequately
judge the true magnitude difference between 20x and 25x communities and
higher cycle numbers analyzed, even though there is a clear trend to
suggest that 20x is very different then 25x {[}Figure 3{]}. This
finding, in conjunction with the PERMANOVA results, suggest that cycle
number can change bacterial community calculations but that these
differences are minimal once 25 cylces are reached. Increasing the
sub-sampling depth, for some DNA polymerase, may reduce some of these
observed community differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence
differences between the different HiFi DNA polymerases {[}Figure 5{]}.
The chimera prevalence was strongly correlated with the number of OTUs
which is relied upon heavily for many community metric calculations.
However, Bray-Curtis analysis with PERMANOVA showed few differences
based on DNA polymerase. It is possible that many of the increased
number of OTUs are not highly abundant allowing the Bray-Curtis index to
be able to successfully downweight these respective OTUs (Minchin,
1987). The choice of downstream diversity metric could be an important
consideration in helping to mitigate these observed changes due to high
chimera prevalence in specific HiFi DNA polymerases (e.g.~Platinum).

Our observations suggest that there are clear HiFi DNA polymerase
dependent differences in both per base error rate and chimeras that
cannot be removed using bioinformatic approaches {[}Figure 4 \& 5{]}.
Although it may be a natural assumption that the variation may be due to
the DNA polymerase family, the highest chimera rate, from Platinum, was
a family A polymerase while the lowest, from Accuprime, was also an A
polymerase (Ishino \& Ishino, 2014). In fact, from the material safety
data sheet (MSDS), the differences between the two mixes is not
immediately apparent. Both Accuprime and Platinum contain a recombinant
\emph{Taq} DNA polymerase, a \emph{Pycrococcus} spp GB-D polymerase and
a platinum \emph{Taq} antibody. It is possible that differences in how
the recombinant \emph{Taq} was generated could be the main reason for
the differences in chimera rate.
